A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2016-01-29 archive/v3/docs_v3/data_v3.02.tex
Cleaned up (files are moved to archive now)

\paragraph{Stellar kinematics.} For the dynamical modelling we use the stellar kinematics for J1331 measured by \citet{SWELLSV}. They obtained long-slit spectra along J1331's major-axis with the Low Resolution Imaging Spectrograph (LRIS) on the Keck I 10m telescope. The width of the slot was $1''$ and the seeing conditions had a FWHM of $\sim 1.1''$. Spectra for spatial bins of different widths along the major axis were extracted. They measured line-of-sight stellar rotation velocities ($v_\text{rot}$) and stellar velocity dispersion ($\sigma$) by fitting Gaussian line profiles to emission \Wilma{[TO DO: Check if really emission lines or, more likely, absorption lines]} lines in these spectra \Wilma{[TO DO: Find out which lines]}. Gas kinematics were extracted from fits to H$\alpha$ and NII lines, as tracers for ionized gas.
\\The stellar kinematics, $v_\text{rot}$, $\sigma$ and $v_\text{rms}^2=v_\text{rot}^2 + \sigma^2$ are shown in Figure \ref{fig:kinematics}. The rotation curve reveals a counter-rotating core within $2''\simeq$ 4 kpc. Outside of $\sim 3.5''$ there is a steep drop in the dispersion, which is expected at the boundary between the pressure supported bulge and the rotationally supported disk, which appears around this radius in the F450W filter in Figure \ref{fig:F450W}. However, in the brighter F814W filter in Figure \ref{fig:F814W}  the large reddish bulge extends out to $\sim5''$. 
\\Inside of $\sim 4''$, the data appears to be symmetric, outside of this the assumption of axisymmetry seems not to be valid anymore, considering the data. We add $-2.3~\text{km s}^{-1}$ to the $v_\text{rot}$ to ensure $v_\text{rot}(R'=0) \sim 0$ as a possible correction term for a misjudgement of the systemic velocity. We also symmetrize the data within $4''$ and asign a minimum error of $\delta v_\text{rms} > 5~\text{km s}^{-1}$ to the $v_\text{rms}$ data. In the JAM modelling, which is based on the assumption of axisymmetry, only stellar kinematics with $R'  \lesssim 2.5''$ or $R' \lesssim 4''$ are used. Another reason to restrict to modelling on the bulge region is that our MGE in Table \ref{tab:MGEF814W} is only a good representation of J1331's F814W light distribution inside $\sim 5''$.

 \Wilma{[TO DO: Maybe mention the following details better in the data section.... "Sufficiently inclined galaxies were picked by eye. Follow-up high resolution imaging with the Hubble Space Telescope's (HST) Wide-Field Planetary Camera 2 (WFPC2) was performed to confirm their lensing status \citep{SWELLSI}."]}

>>>>

\paragraph{Stellar kinematics.} For the dynamical modelling we use the stellar kinematics for J1331 measured by \citet{SWELLSV}. They obtained long-slit spectra along J1331's major-axis with the Low Resolution Imaging Spectrograph (LRIS) on the Keck I 10m telescope. The width of the slit was $1''$ and the seeing conditions had a FWHM of $\sim 1.1''$. 1D Spectra for spatial bins of different widths along the major axis were extracted. They measured line-of-sight stellar rotation velocities ($v_\text{rot}$) and stellar velocity dispersion ($\sigma$) from the spectrum containing the absorption lines Mgb (5177 \AA) and FeII (5270, 5335 and 5406 \AA) (analogous to \citealt{SWELLSII}). Gas kinematics were extracted from Gaussian line profile fits to H$\alpha$ (6563 \AA) and [NII] (6583 and 6548 \AA) emission lines, as tracers for ionized gas.

The stellar kinematics, $v_\text{rot}$, $\sigma$ and $v_\text{rms}^2=v_\text{rot}^2 + \sigma^2$ are shown in Figure \ref{fig:kinematics}. The rotation curve reveals a counter-rotating core within $2''\simeq$ 4 kpc. Outside of $\sim 3.5''$ there is a steep drop in the dispersion. This could indicate the boundary between the pressure supported bulge and the rotationally supported disk, which appears around this radius in the F450W filter in Figure \ref{fig:F450W}. However, in the brighter F814W filter in Figure \ref{fig:F814W} the large reddish bulge extends out to $\sim5''$.

Inside of $\sim 4''$, the data appears to be symmetric, outside of this the assumption of axisymmetry seems not to be valid anymore, considering the data. We add $-2.3~\text{km s}^{-1}$ to the $v_\text{rot}$ to ensure $v_\text{rot}(R'=0) \sim 0$ as a possible correction term for a misjudgement of the systemic velocity. We also symmetrize the data within $4''$ and asign a minimum error of $\delta v_\text{rms} > 5~\text{km s}^{-1}$ to the $v_\text{rms}$ data. In the JAM modelling, which is based on the assumption of axisymmetry, only stellar kinematics with $R'  \lesssim 2.5''$ or $R' \lesssim 4''$ are used. Another reason to restrict to modelling on the bulge region is that our MGE in Table \ref{tab:MGEF814W} is only a good representation of J1331's F814W light distribution inside $\sim 5''$.

### 2016-01-29 archive/v3/docs_v3/discussion_ML_v3.03.tex
Cleaned up (files are moved to archive now)

Some of the ambiguities in recovering J1331's matter distribution could be resolved by learning more about stellar populations with different IMFs in J1331. In particular, a sophisticated guess for the stellar mass-to-light ratio in the bulge could be compared to our very reliable measurement of the total mass-to-light ratio inside the Einstein radius (see Table \ref{tab:einsteinML}). This would then either support or contradict the presence of a significant amount of DM in the bulge.

@@

\citet{SWELLSI} estimated J1331's stellar bulge mass given a Salpeter IMF and measured the I-band AB magnitude of the bulge. Transformed to a stellar I-band mass-to-light ratio, their results would correspond to $\Upsilon_\text{I,*}^\text{sal} = 4.7 \pm 1.2$ (see Table \ref{tab:previousresults}). This is not too far from $\Upsilon_\text{I,*} = 4.2 \pm 0.2$ (see Table \Wilma{[TO DO]}), which we found when including a NFW halo in the JAM modelling.

>>>>

Some of the ambiguities in recovering J1331's matter distribution could be resolved by learning more about stellar populations with different IMFs in J1331. In particular, a sophisticated guess for the stellar mass-to-light ratio in the bulge could be compared to our very reliable measurement of the total mass-to-light ratio inside the Einstein radius $\Upsilon_\text{I,tot}^\text{ein} = 5.56 \Upsilon_{I,\odot}$. This would then either support or contradict the presence of a significant amount of DM in the bulge.

@@

\citet{SWELLSI} estimated J1331's stellar bulge mass given a Salpeter IMF and measured the I-band AB magnitude of the bulge. Transformed to a stellar I-band mass-to-light ratio, their results would correspond to $\Upsilon_\text{I,*}^\text{sal} = 4.7 \pm 1.2$ (see Table \ref{tab:previousresults}). This is not too far from $\Upsilon_\text{I,*} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}), which we found when including a NFW halo in the JAM modelling.

### 2016-01-29 archive/v3/docs_v3/discussion_kinematics_v3.03.tex
Cleaned up (files are moved to archive now)

%==============
\begin{figure}
\centering
  \includegraphics[width=0.9\linewidth]{fig/B4_jam_profiles_errors_short_vcirc.ps}
  \caption{Comparison of the circular velocity curve of J1331's inner regions for different models: (a) JAM model with NFW DM halo from Section \ref{sec:results_JAM_NFW} and Figure \ref{fig:modelB4_models} (green). (b) Lens model from Section \ref{sec:results_lensing} and Figure \ref{fig:JAM_modelL} with $\alpha = 1.1$ (red dashed line), $\alpha = 1$ (red solid line) and $\alpha=0.9$ (red dash-dotted line). (c) Mass-follows-light model, which uses the F814W surface brightness in Table \ref{tab:MGEF814W} and the mass-to-light ratio in the Einstein radius, $\Upsilon^\text{ein}_\text{I,tot} = 5.56$, to generate a mass distribution, as in Figure \ref{fig:lenscomparelight} (orange line).  (d) Model from gas kinematics and Einstein mass found by \citet{SWELLSV} (their Figure 2, best model with 68\% confidence region) (blue lines). \Wilma{[TO DO: In Plot, change symbol for M/L einstein and use full number. Mass follows light curve orange. Dutton blue.]}}
  \label{fig:vcirc_comparison}
\end{figure}
%==============

In Figure \ref{fig:vcirc_comparison} we compare the circular velocity curve found by \citet{SWELLSV} with a mass-follows-light model scaled to fit our Einstein mass (by multiplying the light distribution in Table \Wilma{[TO DO]} with $\Upsilon_\text{I,tot}^\text{ein} = 5.6$). Within $R < 5''$ they agree with each other.

The models in this work used more than just the Einstein mass to constrain the matter distribution at small radii: The lens mass model constrained also the shape of the mass distribution within the lensing image configuration at $R_\text{ein} \sim 1''$. The dynamical models used stellar kinematics inside $R' \simeq 3.5''$ \Wilma{[TO DO: Check]}. We compare the lens mass model's $v_\text{circ}$ (for $\alpha=1.0\pm 0.1$) with the NFW JAM model (Table \Wilma{[TO DO]}) in Figure \ref{fig:vcirc_comparison} as well. Within and around $R_\text{ein}$ they are consistent with each other, even though they where independently derived. They do not agree with the result by \citet{SWELLSV} and in Section \ref{sec:results_JAM_SB} we showed, that ``mass-follows-light'' is not a good model for J1331.

>>>>

%==============
\begin{figure}
\centering
  \includegraphics[width=0.9\linewidth]{fig/B4_jam_profiles_errors_short_vcirc.ps}
  \caption{Comparison of the circular velocity curve of J1331's inner regions for different models: (a) JAM model with NFW DM halo from Section \ref{sec:results_JAM_NFW} and Figure \ref{fig:modelB4_models} (green). (b) Lens model from Section \ref{sec:results_lensing} and Figure \ref{fig:JAM_modelL} with $\alpha = 1.1$ (red dashed line), $\alpha = 1$ (red solid line) and $\alpha=0.9$ (red dash-dotted line). (c) Mass-follows-light model, which uses the F814W surface brightness in Table \ref{tab:MGEF814W} and the mass-to-light ratio in the Einstein radius, $\Upsilon^\text{ein}_\text{I,tot} = 5.56$, to generate a mass distribution, as in Figure \ref{fig:lenscomparelight} (orange line).  (d) Model from gas kinematics and Einstein mass found by \citet{SWELLSV} (their Figure 2, best model with 68\% confidence region) (blue lines).}
  \label{fig:vcirc_comparison}
\end{figure}
%==============

In Figure \ref{fig:vcirc_comparison} we compare the circular velocity curve found by \citet{SWELLSV} with a mass-follows-light model scaled to fit our Einstein mass (by multiplying the light distribution in Table \ref{tab:MGEF814W} with $\Upsilon_\text{I,tot}^\text{ein} = 5.56$). Within $R < 5''$ they agree with each other.

The models in this work used more than just the Einstein mass to constrain the matter distribution at small radii: The lens mass model constrained also the shape of the mass distribution within the lensing image configuration at $R_\text{ein} \sim 1''$. The dynamical models used stellar kinematics inside $R' \simeq 3.5''$ \Wilma{[TO DO: Check]}. We compare the lens mass model's $v_\text{circ}$ (for $\alpha=1.0\pm 0.1$) with the NFW JAM model (Table \ref{tab:modelB4_bestfit}) in Figure \ref{fig:vcirc_comparison} as well. Within and around $R_\text{ein}$ they are consistent with each other, even though they where independently derived. They do not agree with the result by \citet{SWELLSV} and in Section \ref{sec:results_JAM_SB} we showed, that ``mass-follows-light'' is not a good model for J1331.
